The Operational Space of International Civil Society Under Authoritarian Regulations: The Case of China

### 2023-12-04 manuscript/manuscript.qmd
Clean and shrink data and methods section

[^ngo-clarification]: "Civil society" is a broad concept that includes a range of organizational forms, both formal and informal. In this paper, we look at nongovernmental organizations (NGOs), often called nonprofit organizations or third sector organizations—formal non-profit-seeking organizations that are neither privately-owned nor government-controlled. We define international NGOs as organizations that conduct work predominantly in at least one other country outside of their home country.

@@

By focusing on INGO operational space in China, we provide a unique view into the *de facto* regulatory environment for foreign civil society in an authoritarian state. Many previous studies use broad indexes of civil society openness, such as the core civil society index from the Varieties of Democracy project [@LianMurdie:2023; @ChaudhryHeiss:2022]. These measures capture the general regulatory environment for all civil society—including domestic NGOs, INGOs, and informal organizations—but do not reflect the specific implementation of laws for INGOs. In this paper, though, we use a mixed methods approach to explore the regulatory environment for INGOs in particular, providing much a richer description of authoritarian regulation of international civil society.

Below, we provide a brief overview of China's 2017 ONGO Law and the different regulatory mechanisms—and practical limitations—it provides for INGOs to obtain legal access to the country. We then summarize existing research on authoritarian regulation of INGOs and the differential application of these laws based on regime preferences and organizational characteristics, and propose a set of empirical expectations based on this literature. We test these expectations with a mixed methods research design, combining Bayeisan analysis of administrative data from 633 formally registered INGOs with a comparative case study of Greenpeace and The Nature Conservancy, two INGOs working on similar issue areas which took different approaches to operating in China after 2017. We conclude with a broader discussion of authoritarian civil society regulations and propose future methods for exploring the practical operating environment for INGOs working in such regimes.

@@

The 2017 ONGO Law's provision of two legal pathways for registration—each with different degrees of freedom—is an example of the broader phenomenon of closing civic space. In the past two decades, authoritarian states have used administrative and legal measures to control and curtail the influence of global civil society. A rich literature in comparative politics explores how authoritarian regimes counterintuitively use democratic institutions for their own benefit and survival [@Kendall-TaylorFrantz:2014a; @GandhiPrzeworski:2007]. Authoritarian states hold elections [@LevitskyWay:2010], permit popular protests [@Weiss:2013], and allow opposition parties in the legislature and set executive term limits [@Meng:2020]. However, autocrats use these democratic-appearing institutions to limit possible threats to regime stability. Elections, protests, and apparent political reform are allowed, but only in ways that bolster the government's stability, and these institutions are kept weak and "dependent on the regime to ensure that they do not develop any real power or autonomy" [@FrantzEzrow:2011, 7].

@@

These laws, however, are a blunt tool that tend to target civil society as a whole—authoritarian states do not create one set of formal laws that applies only to regime-friendly INGOs and another set that applies to more risky INGOs. Instead, there is a divergence between the formal *de jure* laws and the *de facto* implementation of their accompanying regulations. In practice, NGO laws tend to not be applied universally and formal laws often serve as a "weapon hanging on the wall that never fires," acting a warning to potentially uncooperative NGOs [@Kozenko:2015]. Instead, autocrats engage with INGOs in opposing ways, providing financial support and opportunities for collaboration with those that align with government policy preferences, and using legal mechanisms to restrict or expel those that do not [@ToeplerZimmerFrohlich:2020; @Plantan:2022; @LiFarid:2023].

@@

In @fig-theory, we propose that the interplay between INGO characteristics and government preferences is a key determinant of individual INGO operating space [@LiFarid:2023; @ChaudhryHeiss:2021]. Authoritarian states seek out INGOs that provide direct benefits and public goods, while avoiding INGOs that threaten their stability. For instance, states in need of foreign aid and development will be more open to INGOs working on those issues, while states looking for increased trade will seek out INGOs focused on economic partnerships. Because their primary issue areas are aligned with regime preferences, INGOs benefit from greater operational space and face fewer restrictions. In contrast, states engaged in human rights abuses or fearful of foreign influence over other domestic policies will reduce—or even close—the operational space allowed for INGOs addressing these issues. Notably, the arrow between government preferences and issue area is bidirectional, reflecting the government's ability to use civil society regulations to directly realign INGO programming in the regime's favor. Finally, the natural scope of an INGO's activities can shape its allowed operational space outside of authoritarian stability seeking calculus. Many INGOs purposely keep themselves small, non-confrontational, and non-political and limit themselves to work in only a small part of their target countries, having no plans to go beyond their hyperlocal scope. This naturally limits their operational space, but these limits are self-imposed and not directly related to government regulations.

@@

We test these theoretical expectations using a mixed methods approach. Using

1: issue - stats + case study 2: localness - stats 3: timing - stats

-   Stats + case study: **Local**
-   Stats + case study: **Contentiousness**
-   Case study: **Relationship with PSU and government** (stats/descriptive bc lots of chambers of commerce)

@@

TODO Andrew: Make it clear that this section is just about the data for the statsy stuff, not the case study

The data used in this paper was collected from INGOs' public registration documents disclosed by the Ministry of Public Security's Overseas NGO Office online service platform. The platform is a website maintained by the registrar of INGOs and regularly publicize the administrative documentation of all the ROs and temporary activities registered or filled across China. The data used in this study spans from January 1, 2017, when the law took effect, to December 31, 2021, with information for all 633 registered organizations approved as of the cutoff point. The filing documents for ROs disclose the name, the address, the registration date, the registration authority (which public security bureau regulates it), the PSU, the issue area or purposes of the RO, and the specific provinces they can work in. We merge our INGO data with additional data of Chinese INGO data from ChinaFile, a project that translate and code the MPS platform data, to include their English organization names, for the sake of transparency and accessibility.

We code the variables we use in the empirical analysis from the registration information. Most of our data cleaning work is an objective data wrangling process, such as parsing out the province each RO registered in from the address, or calculating the total number provinces approved in the registration document. Meanwhile, the qualitative coding of issue areas is conducted by the authors, based on consistent coding rules agreed by the authors.

@@

### Dependent variable

As explained, the dependent variable is the count of provinces that INGOs are authorized to work in as documented in their official registration record. The documentation lists the specific provincial regions each INGO RO can conduct activities in and we derive the number of provinces based on the lists. Since the maximum count of provinces INGOs can possibly work in is 32, this variable ranges from 1 to 32.

### Independent variables

Our key independent variables correspond to each of our three empirical expectations. First we code each INGO's issue area based on its mission statement reported in its registration documentation. Given there is certain level of subjectivity in the coding process, the coding of issue areas are cross-checked between the researchers---Author 1 assigned each organization into one of the issue areas defined in Article 3 of the ONGO Law, and Author 2 cross-checked the coding for ambiguous cases. We include our coding rules for overlapping issue areas in the appendix (TODO). We categorize each organization into nine different issue areas (see @tbl-issues).

@@

Our second independent variable is a dichotomous indicator of whether the INGO explicitly states that it intends to operate locally (below province level). There are cases where the aim of the INGO is attached to a locality, either having a county or city name in their RO name or purposes, or the INGO are actually set up by people in China and they registered the RO to communicate between the "hometown" and the INGO set up abroad. We thus include this variable in our analysis to explore how having local connections is associated with the operational space of ROs.

Our final independent variable measures how much time elapsed between the January 1, 2017 (when the law took effect) and the organization's official registration date. There is great variance in the timing of registering an RO among INGOs, since there has been a *de facto* grace period for INGOs to seek legal presence after the law took effect [@Shieh:2018]. This is largely because both INGOs and the regulators needed some transitional time to explore the specific implementation of the law as large ambiguity was left when the law was passed. We use this temporal variable as a proxy for the motivation or eagerness of INGOs to register an RO, for some INGOs are inclined to avoid the cumbersome registration procedures at the early stage if there is no real consequences for continuing operations without registration, while others took the time to talk with multiple potential PSUs to find a more suitable superior or regulatory dynamic. (TODO: find a citation for this, probably Spires, if there is one.)

@@

## Descriptive overview of data

@@

Following our first empirical expectation, we predict that INGOs working on less contentious issues should be registered in more provinces and have more geographic flexibility in operating throughout the country. To explore this, we can take advantage of the simultaneous continuous and discrete features of ordered Beta regression and see how often an INGO in each issue area is predicted to work in a single province, across all provinces, or some number in between. @fig-issue-preds(a) presents a summary of predicted posterior outcomes for 1,000 hypothetical INGOs, varying each possible issue area and holding all other variables at their mean or modal values. Several notable trends emerge from the model results. INGOs focused on arts and culture are overwhelming the most likely kind of organization to work nationally---arts and culture organizations are predicted to work in all 32 provinces 80% of the time (with a credible interval of 78--83%) and are only predicted to work in a single province a rare 3--5% of the time. Education-focused INGOs, on the other hand, are the least likely kind of organization to be registered in all 32 provinces and are predicted to do so only 32--38% of the time. Education INGOs are the most likely to be registered in only one province, with that predicted outcome appearing 24--29% of the time---at least 10 percentage points higher than any other issue area's predicted proportion. All other issue areas follow roughly the patterns in predictions: the most common outcome is for organizations to work nationally (50--60% of the time), followed by working in at least 2 provinces (25--40% of the time), followed by working in a single province (10--20% of the time). @fig-issue-preds(b) shows the continuous posterior distribution of the average number of provinces INGOs are expected to be registered in across all types of issue areas, but accounting for the 1- and 32-only provinces simultaneously. We can see a similar trend---arts and culture INGOs are predicted to work in 28--30 provinces on average, while education INGOs are only predicted to work in 12--20. All other issues range between 20--28 provinces.

>>>>

[^ngo-clarification]: "Civil society" is a broad concept that includes a range of organizational forms, both formal and informal. In this paper, we look at nongovernmental organizations (NGOs), often called nonprofit organizations or third sector organizations---formal non-profit-seeking organizations that are neither privately-owned nor government-controlled. We define international NGOs as organizations that conduct work predominantly in at least one other country outside of their home country.

@@

By focusing on INGO operational space in China, we provide a unique view into the *de facto* regulatory environment for foreign civil society in an authoritarian state. Many previous studies use broad indexes of civil society openness, such as the core civil society index from the Varieties of Democracy project [@LianMurdie:2023; @ChaudhryHeiss:2022]. These measures capture the general regulatory environment for all civil society---including domestic NGOs, INGOs, and informal organizations---but do not reflect the specific implementation of laws for INGOs. In this paper, though, we use a mixed methods approach to explore the regulatory environment for INGOs in particular, providing much a richer description of authoritarian regulation of international civil society.

Below, we provide a brief overview of China's 2017 ONGO Law and the different regulatory mechanisms---and practical limitations---it provides for INGOs to obtain legal access to the country. We then summarize existing research on authoritarian regulation of INGOs and the differential application of these laws based on regime preferences and organizational characteristics, and propose a set of empirical expectations based on this literature. We test these expectations with a mixed methods research design, combining Bayeisan analysis of administrative data from 633 formally registered INGOs with a comparative case study of Greenpeace and The Nature Conservancy, two INGOs working on similar issue areas which took different approaches to operating in China after 2017. We conclude with a broader discussion of authoritarian civil society regulations and propose future methods for exploring the practical operating environment for INGOs working in such regimes.

@@

The 2017 ONGO Law's provision of two legal pathways for registration---each with different degrees of freedom---is an example of the broader phenomenon of closing civic space. In the past two decades, authoritarian states have used administrative and legal measures to control and curtail the influence of global civil society. A rich literature in comparative politics explores how authoritarian regimes counterintuitively use democratic institutions for their own benefit and survival [@Kendall-TaylorFrantz:2014a; @GandhiPrzeworski:2007]. Authoritarian states hold elections [@LevitskyWay:2010], permit popular protests [@Weiss:2013], and allow opposition parties in the legislature and set executive term limits [@Meng:2020]. However, autocrats use these democratic-appearing institutions to limit possible threats to regime stability. Elections, protests, and apparent political reform are allowed, but only in ways that bolster the government's stability, and these institutions are kept weak and "dependent on the regime to ensure that they do not develop any real power or autonomy" [@FrantzEzrow:2011, 7].

@@

These laws, however, are a blunt tool that tend to target civil society as a whole---authoritarian states do not create one set of formal laws that applies only to regime-friendly INGOs and another set that applies to more risky INGOs. Instead, there is a divergence between the formal *de jure* laws and the *de facto* implementation of their accompanying regulations. In practice, NGO laws tend to not be applied universally and formal laws often serve as a "weapon hanging on the wall that never fires," acting a warning to potentially uncooperative NGOs [@Kozenko:2015]. Instead, autocrats engage with INGOs in opposing ways, providing financial support and opportunities for collaboration with those that align with government policy preferences, and using legal mechanisms to restrict or expel those that do not [@ToeplerZimmerFrohlich:2020; @Plantan:2022; @LiFarid:2023].

@@

In @fig-theory, we propose that the interplay between INGO characteristics and government preferences is a key determinant of individual INGO operating space [@LiFarid:2023; @ChaudhryHeiss:2021]. Authoritarian states seek out INGOs that provide direct benefits and public goods, while avoiding INGOs that threaten their stability. For instance, states in need of foreign aid and development will be more open to INGOs working on those issues, while states looking for increased trade will seek out INGOs focused on economic partnerships. Because their primary issue areas are aligned with regime preferences, INGOs benefit from greater operational space and face fewer restrictions. In contrast, states engaged in human rights abuses or fearful of foreign influence over other domestic policies will reduce---or even close---the operational space allowed for INGOs addressing these issues. Notably, the arrow between government preferences and issue area is bidirectional, reflecting the government's ability to use civil society regulations to directly realign INGO programming in the regime's favor. Finally, the natural scope of an INGO's activities can shape its allowed operational space outside of authoritarian stability seeking calculus. Many INGOs purposely keep themselves small, non-confrontational, and non-political and limit themselves to work in only a small part of their target countries, having no plans to go beyond their hyperlocal scope. This naturally limits their operational space, but these limits are self-imposed and not directly related to government regulations.

@@

# Data and methods

We examine these theoretical expectations using a mixed methods approach. First, we explore a rich dataset of all ROs registered in China between 2017–2021 using a combination of descriptive statistics and Bayesian models to understand how the international nonprofit sector in China has been reshaped since the implementation of the ONGO law. We then complement this statistical analysis with a comparative case study of two INGOs---Greenpeace and The Nature Conservancy---to explore in detail how the ONGO has specifically influenced organizational operational space.

@@

Our data comes from INGOs' public registration documents disclosed by the Ministry of Public Security's Overseas NGO Office, which maintains a website with complete documentation of all the ROs and temporary activities registered or filled across China. We include all INGO registrations between January 1, 2017, when the law took effect, to December 31, 2021, with information for all `r nrow(ongo)` registered organizations. The filing documents for ROs disclose each INGO's name, address, registration date, regulating Public Security Bureau, professional supervisory unit (PSU),  issue area or purposes of the RO, and the specific provinces they can work in. We merge this INGO data with translated English organization names data from ChinaFile. Our data and reproducible code is available at `ANONYMIZED_DOI_URL`.

@@

Our dependent variable is the count of provinces that INGOs are authorized to work in. We derive this value from the official registration record, which lists all the provincial regions each INGO RO can conduct activities in. This value ranges from 1–32 (the total number of Chinese provinces).

Our key independent variables correspond to each of our three theoretical expectations. First we code each INGO's issue area based on its mission statement reported in its registration documentation. Given there is certain level of subjectivity in the coding process, we cross-checked our manual coding of issue areas---Author 1 assigned each organization into one of the issue areas defined in Article 3 of the ONGO Law, and Author 2 cross-checked the coding for ambiguous cases. We include our coding rules for overlapping issue areas in the appendix (TODO). We categorize each organization into nine different issue areas (see @tbl-issues).

@@

Our second independent variable is a dichotomous indicator of whether the INGO explicitly states that it intends to operate locally or subprovincially. There are cases where the aim of the INGO is attached to a locality, either having a county or city name in their RO name or purposes, or the INGO is actually set up by people in China who registered the RO to communicate between the "hometown" and the INGO set up abroad.

Our final independent variable measures how much time elapsed between January 1, 2017 (when the law took effect) and the organization's official registration date. There is great variance in the timing of registering an RO among INGOs, since there has been a *de facto* grace period for INGOs to seek legal presence after the law took effect [@Shieh:2018]. This is largely because both INGOs and the regulators needed some transitional time to explore the specific implementation of the law as large ambiguity was left when the law was passed. We use this temporal variable as a proxy for the motivation or eagerness of INGOs to register an RO, for some INGOs are inclined to avoid the cumbersome registration procedures at the early stage if there is no real consequences for continuing operations without registration, while others took the time to talk with multiple potential PSUs to find a more suitable superior or regulatory dynamic. (TODO: find a citation for this, probably Spires, if there is one.)

@@

## Descriptive overview

@@

Following our first empirical expectation, we predict that INGOs working on less contentious issues should be registered in more provinces and have more geographic flexibility to work throughout the country. To explore this, we can take advantage of the simultaneous continuous and discrete features of ordered Beta regression and see how often an INGO in each issue area is predicted to work in a single province, across all provinces, or some number in between. @fig-issue-preds(a) presents a summary of predicted posterior outcomes for 1,000 hypothetical INGOs, varying each possible issue area and holding all other variables at their mean or modal values. Several notable trends emerge from the model results. INGOs focused on arts and culture are overwhelming the most likely kind of organization to work nationally---arts and culture organizations are predicted to work in all 32 provinces 80% of the time (with a credible interval of 78--83%) and are only predicted to work in a single province a rare 3--5% of the time. Education-focused INGOs, on the other hand, are the least likely kind of organization to be registered in all 32 provinces and are predicted to do so only 32--38% of the time. Education INGOs are the most likely to be registered in only one province, with that predicted outcome appearing 24--29% of the time---at least 10 percentage points higher than any other issue area's predicted proportion. All other issue areas follow roughly the patterns in predictions: the most common outcome is for organizations to work nationally (50--60% of the time), followed by working in at least 2 provinces (25--40% of the time), followed by working in a single province (10--20% of the time). @fig-issue-preds(b) shows the continuous posterior distribution of the average number of provinces INGOs are expected to be registered in across all types of issue areas, but accounting for the 1- and 32-only provinces simultaneously. We can see a similar trend---arts and culture INGOs are predicted to work in 28--30 provinces on average, while education INGOs are only predicted to work in 12--20. All other issues range between 20--28 provinces.
